### 2021-11-26 plos_latex_template.tex
Few abstract/pillars changes ~ @jananiravi

The authors of this article participated in the organization of the annual conference of the R Project for Statistical Computing, held in July 2021. useR! conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we outline our most important lessons learned in 10 simple rules to host an inclusive conference.\fede{general comment on this: we should clearly state that this whole-virtual setting was a clear consequence of the pandemics, AND it was planned like this from the beginning, and not changing in-between to this format} These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level. The rules were learned during a virtual conference, but serve also as guidelines to in-person and hybrid events.

@@

To make a more inclusive conference, it is necessary to identify which groups of people are being excluded, the barriers they may face to be part of the conference, and strategies to overcome them. 
The first rule is about embracing all dimensions of diversity and prioritizing the
 participation of groups that are often marginalized.
However, increasing diversity is not enough. Rule 2 is about creating a safe and welcoming environment to make everyone feel part of the conference and the community. 
Inclusion should start from the organizing team. Rule 3 refers to the importance of starting with an inclusive and diverse organizing team, and provides advice on work dynamics.

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (e.g., cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built. \as{People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia. No form of oppression operates in isolation from the others, which means some groups are particularly vulnerable and systematically left behind.}
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}. 
Recognizing this systematic exclusion, particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against.
These are the groups we need to make more effort to include.
Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
% vision
% back to the indicators? 
% realistic but bold (back?)
% be aware and promote awareness
%¡LAS BARRERAS!
%Express the diversity and inclusion vision for your conference by writing a diversity statement.

@@

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
%in 2021 an inclusive conference has to be an online or hybrid conference. this means a radical change in the functioning of all spaces, to include people who are at home for any reason. acá contexto de covid - within ... y ser más firmes con lo de hybrid. joo - wenger. 
%primer parrafo como está cite joo, cosas prácticas de online - están en las citas, poner acá. 2. networking hybrid qué es, tipos, active and distributed serían nuestras opciones - omitir las otras por no ser inclusivas. todo lo inclusivo de online con todo lo inclusivo de presencial. un par de ejemplos de lo presencial. o adaptaciones que requiere la parte presencial cuando se vuelve híbrido (chat, q and a, streaming) opportunities of networking - que ser chair o keynote o presenter no cambie si estás online o presencial, no requiera que estés presente. 
%practical: choose the platforms and be aware that reach varies between countries
As stated in [the nature correspondence]: The COVID pandemic forced us to embrace online conferences (see C. Woolston Nature 582, 135–136; 2020), which remove some of the barriers that disproportionately affect marginalized groups. These include the cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications (H. J. Niner \& S. N. Wassermann Front. Mar. Sci. 8, 638025; 2021; Gewin, V. Nature 569, 297–299; 2019). Inclusive conferences must have an online component [cite all the other references].
In-person interaction at conferences is priceless but often this is true and possible only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of attendees but in terms of speakers  \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several time zones or deciding to rotate the favored time zone every year, without depending on a conference central location.

>>>>

The authors of this article recently organized useR! 2021, the annual conference of the R Project for Statistical Computing. useR! conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference providing a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we outline our most important lessons learned in 10 simple rules to host an inclusive conference. These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules also apply to regional events. These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events as we exit the pandemic.

@@

To organize a more inclusive conference, it is necessary to identify upfront which groups of people are being excluded, the barriers that keep them from active partaking, and targeted strategies to overcome them. 
\textbf{Rule 1} is about embracing all dimensions of diversity and prioritizing the participation of groups that are often marginalized.
However, increasing diversity alone is not enough. \textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the conference attendees and the larger community. 
Inclusion should start from the organizing team. \textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team, and provides tips on work dynamics.

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (\textit{e.g.}, cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built. \as{People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia. No form of oppression operates in isolation from the others, which means some groups are particularly vulnerable and systematically left behind.}
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}. 
Recognizing this systematic exclusion, particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against.
These are the groups we need to make more effort to include.
Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
% vision
% back to the indicators? 
% realistic but bold (back?)
% be aware and promote awareness
%¡LAS BARRERAS!
%Express the diversity and inclusion vision for your conference by writing a diversity statement.

@@

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
%1. in 2021 an inclusive conference has to be an online or hybrid conference. this means a radical change in the functioning of all spaces, to include people who are online for any reason. we have elements to advocate for hybrid and lots of refs about online. cite joo and online refs. 
%practical: choose the platforms and be aware that reach varies between countries. check accessibility. understand bandwidth limitations. online does not make it automatically inclusive.
%2. hybrid. we advocate for active hybrid and distributed hybrid conferences because it's the most inclusive formats (not passive hybrid or semi-passive hybrid). we want all the inclusiveness of the in-person component with the _practical_ adaptations for hybrid (chat platform, q and a, streaming and networking) being online or in-person should not affect the possibility to be a chair a keynote or a presenter. no need to be present. a potential problem: opportunities of networking.
As stated in [the nature correspondence]: The COVID pandemic forced us to embrace online conferences (see C. Woolston Nature 582, 135–136; 2020), which remove some of the barriers that disproportionately affect marginalized groups. These include the cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications (H. J. Niner \& S. N. Wassermann Front. Mar. Sci. 8, 638025; 2021; Gewin, V. Nature 569, 297–299; 2019). Inclusive conferences must have an online component [cite all the other references].
In-person interaction at conferences is priceless but often this is true and possible only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of attendees but in terms of speakers  \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several time zones or deciding to rotate the favored time zone every year, without depending on a conference central location.
